Estimating the genome-wide contribution of selection to temporal allele frequency change

### 2020-05-10 manuscript/manuscript.tex
editing Ne section

 Rapid phenotypic adaptation is often observed in natural populations and
 selection experiments.  However, detecting the genome-wide impact of this
 selection is difficult, since adaptation often proceeds from standing
 variation and selection on highly polygenic traits, both of which may leave
 faint genomic signals indistinguishable from a noisy background of genetic
 drift.  One promising signal comes from the genome-wide covariance between
 allele frequency changes observable from temporal genomic data, e.g.
 evolve-and-resequence studies. These temporal covariances reflect how the
 change \gc{in allele} frequency at one timepoint is predictive of the
 changes at later timepoints when there is heritable fitness variation in the
 population, \gc{as alleles} can remain associated with selected alleles
 over time \gc{and so be indirectly selected for}. Since genetic drift does not lead to temporal covariance, we can
 use these covariances to estimate what fraction of the variation in allele
 frequency change through time is driven by linked selection. Here, we
 reanalyze \vb{three selection experiments}, \vbout{two \emph{Drosophila
   simulans} evolve-and-resequence studies, and one artificial selection
 experiment in mice}, to quantify the effects of linked selection over short
 timescales using covariance among time-points and across replicates. We
 estimate that at least 17\% to 37\% of allele frequency change is driven by
 selection in these experiments.  Against this background of positive
 genome-wide temporal covariances we also identify signals of negative temporal
 covariance corresponding to reversals in the direction of selection for a
 reasonable proportion of loci over the time course of a selection experiment.
 Overall, we find that in the three studies we analyzed, linked selection has a
 large impact on short-term allele frequency dynamics that is readily
 distinguishable from genetic drift.

@@

A long-standing problem in evolutionary genetics is quantifying the roles of
genetic drift and selection in shaping genome-wide allele frequency changes.
Selection can affect allele frequencies, \vb{both directly and indirectly,}
with the indirect effect coming from the action of selection on \vb{linked}
loci elsewhere in genome e.g. linked selection (\cite{Maynard_Smith1974-lc},
\cite{Charlesworth1993-gb,Nordborg1996-nq}; see \cite{Barton2000-zg} for a
review). Previous work \vbout{on this question} has mostly focused on teasing
apart the impacts of drift and selection on genome-wide diversity using
population samples from a single contemporary timepoint, often by modeling the
correlation between regional recombination rate, gene density, and diversity
created in the presence of linked selection
\parencite{Cutter2013-ba,Sella2009-nx}. This approach has shown linked
selection has a major role in shaping patterns of genome-wide diversity across
the genomes of a range of sexual species
\parencite{Macpherson2007-qt,Andolfatto2007-uy,Begun2007-bg,Beissinger2016-cm,Sattath2011-dr,Williamson2014-oy,Andersen2012-bj,Cutter2010-gi,Elyashiv2016-vt},
and has allowed us to quantify the relative influence of positive selection
(hitchhiking) and negative selection (background selection;
\cite{Nordborg2005-dc,McVicker2009-ax,Andolfatto2007-uy,Macpherson2007-qt,Hernandez2011-gs,Elyashiv2016-vt}).
However, we lack an understanding of both how linked selection acts over
\vb{short time intervals and its full impact on genome-wide allele frequency
changes.}

There are numerous examples of rapid phenotypic adaptation
\parencite{Grant2011-wk,Grant2006-hj,Reznick1997-mh,Franks2007-dr} and rapid,
selection-driven genomic evolution in asexual populations
\parencite{Good2017-om,Bennett1990-bc,Baym2016-kh}.  Yet the polygenic nature
of fitness makes detecting the impact of selection on genome-wide variation
over short timescales in sexual populations remarkably difficult
\vb{\parencite{Latta1998-me, Pritchard2010-tk,Kemper2014-bx}}. This is because
the effect of selection on a polygenic trait (such as fitness) is distributed
across \vb{numerous} loci. This can lead to subtle allele frequency shifts on
standing variation that are difficult to distinguish from background levels of
genetic drift and sampling variance. \vb{Increasingly}, genomic experimental
evolution studies with multiple timepoints, and in some cases multiple
replicate populations, are being used to detect large effect selected loci
\parencite{Turner2011-sx,Turner2012-bm} and differentiate modes of selection
\parencite{Burke2010-tz,Barghi2019-qy,Therkildsen2019-zy}.  In addition these
temporal-genomic studies have begun in wild populations, some with the goal of
finding variants that exhibit frequency changes consistent with fluctuating
selection \parencite{Bergland2014-ij,Machado2018-cs}. In a previous paper, we
proposed that one useful signal for understanding the genome-wide impact of
polygenic linked selection detectable from genomic studies with multiple
timepoints is the temporal autocovariance (\vb{i.e. covariance between two
timepoints}) in allele frequency changes \parencite{Buffalo2019-io}.  These
covariances \vb{can be estimated} from temporal genomic data and are created
when the loci that underly heritable fitness variation perturb the frequencies
of linked \gc{alleles}; in contrast, when genetic drift acts alone in a
closed population, these covariances are zero in expectation \gc{for
  neutral alleles}.  Mathematically,
temporal covariances are useful because it is natural to decompose the total
variance in allele frequency change across a set of time intervals into the
variances and covariances in allele frequency change among time intervals.
Furthermore, biologically, these covariances reflect the extent to which
\gc{allele} frequency changes in one generation predict changes in another
due to a shared selection pressures and associations to selected loci.

@@

We first analyzed \textcite{Barghi2019-qy}, an evolve-and-resequence study with
ten replicate populations exposed to a high temperature lab environment and
evolved for 60 generations, and sequenced every ten generations. Using the
seven timepoints and ten replicate populations, we estimated the genome-wide $6
\times 6$ temporal covariance matrix $\mathbf{Q}$ for each of the ten
replicates. Each row of these matrices represent the temporal covariance
$\cov(\Delta_{_{10}} p_s, \Delta_{_{10}} p_t)$, between the allele frequency
change (in ten-generation intervals, denoted $\Delta_{_{10}} p_t$) in some
initial reference generation $s$ (the row of the matrix), and some later
timepoint $t$ (the column of the matrix). We corrected these matrices for
biases created due to sampling noise, and normalize the entries for
heterozygosity (see Supplementary Materials Sections
\ref{supp:ind-depth-var-corr} and \ref{supp:matrix-correction}). These
covariances are expected to be zero when only drift is acting, as only
heritable variation for fitness can create covariance between allele frequency
changes in a closed population \parencite{Buffalo2019-io}.  Averaging across
the ten replicate temporal covariances matrices, we find temporal covariances
that are statistically significant (95\% block bootstraps CIs do not contain
zero), consistent with linked selection perturbing genome-wide allele frequency
changes over very short time periods. The covariances between all adjacent time
intervals are positive and then decay towards zero as we look at more distant
time intervals (Figure \ref{fig:figure-1} A), as expected when directional
selection affects linked variants' frequency trajectories until ultimately
linkage disequilibrium and the \gc{associated} additive genetic
variance for fitness decays (\vb{which could occur as a population reaches
a new optimum}) \parencite{Buffalo2019-io}. The temporal covariances per
replicate are noisier but this general pattern holds; see Supplementary Figure
\ref{suppfig:barghi-cov-panels}.  \textcite{Barghi2019-qy}'s design means that
the covariances we see in adjacent time intervals are on average ten
generations apart, and given the temporal decay in covariance we see, the
covariances on shorter time-scales (e.g. if adjacent generations had been
sequenced) may well be higher yet (see Supplementary Material Section
\ref{supp:barghi-covs} for more details).

@@

which estimates the \vb{impact of selection} on allele frequency change between
the initial generation $0$ and some later generation $t$, which can be varied
to see how this quantity grows through time. \vb{This measure can intuitively
  be understood as the relative fraction of allele frequency change normally
  thought of as ``drift" that is actually due to selection; the denominator of
  $G(t)$ is $\var(p_t - p_0)$, which under drift alone is proportional to
  $\nicefrac{1}{2 N_e}$. By measuring the contribution of temporal covariances
caused by selection, $G(t)$ looks at the fraction selection contributes to what
looks like a faster rate of drift, or lower $N_e$.} Since
\textcite{Barghi2019-qy} experiment is sequenced every ten generations, in the
numerator for the covariance we use the allele frequency changes between
adjacent timepoints, which are ten generations apart; \vb{thus the strong
  covariances between adjacent generations are not observable and do not
  contribute to the numerator of $G(t)$}. \vb{Additionally, selection also
  inflates the variance in allele frequency change per generation; however,
  this effect cannot be easily distinguished from drift.} \vb{For both these
reasons, our measure $G(t)$ is strongly conservative} (\vb{we demonstrate this
through simulations in Supplementary Material Section \ref{supp:tempblocks}}).
Still, we find a remarkably strong signal. Greater than $20\%$ of total,
genome-wide allele frequency change over 60 generations is the result of
selection (Figure \ref{fig:figure-1} B).

@@

The replicate design of \textcite{Barghi2019-qy} allows us to quantify another
covariance: the covariance in allele frequency change between replicate
populations experiencing convergent selection pressures. These
between-replicate covariances are created in the same way as temporal
covariances: \gc{alleles} linked to a particular fitness background are
expected to have allele frequency changes in the same direction if the
selection pressures are similar. Intuitively, where temporal covariances
reflect that \gc{alleles} associated with heritable fitness backgrounds are
predictive of frequency changes between generations, replicate covariances
reflect that heritable fitness backgrounds common to each replicate predict
(under the same selection pressures) frequency changes between replicates. We
measure this through a statistic similar to a correlation, which we call the
convergent correlation: the ratio of average between-replicate covariance
across all pairs to the average standard deviation across all pairs of
replicates,

@@

Our primary focus here has been on evolution in laboratory populations. It is
unclear whether we should expect a similar impact of selection in natural
populations. In some of these experiments, selection pressures may have been
stronger or more sustained \vb{than} in natural populations
\parencite{Hendry1999-zu,Hairston2005-ga}. Conversely, these lab populations
\gc{were maintained at relatively small census sizes (Table
\ref{table:study_summary})}, which will amplify the role of genetic drift. The
advantage of lab experiments is that they are closed populations\vb{;} in
natural populations temporal covariance could also arise from the systematic
migration of alleles from differentiated populations. Adapting these methods to
natural populations will require either populations that are reasonably closed
to migration, or for the effect of migration to be accounted for possibly
either by knowledge of allele frequencies in source populations or the
identification of migrant individuals.

@@

To remove systematic covariances in allele frequency change caused by tracking
the reference or minor allele, we randomly choose which allele's frequency to
track for each locus. Then, we calculate the variance-covariance matrix of
allele frequency changes using a Python software package we have written,
available at \url{http://github.com/vsbuffalo/cvtk}. This simultaneously
calculates temporal variances and covariances, and replicate covariances and
uses the sampling depth and number of diploid individuals to correct for bias
in the variance estimates and a bias that occurs in covariance estimates
between adjacent timepoints due to shared sampling noise (see Supplementary
Material Sections \ref{supp:ind-depth-var-corr}, \ref{supp:cov-corr}, and
\ref{supp:matrix-correction} for mathematical details of these estimators). We
assess that our bias correction procedure is working adequately through a
series of diagnostic plots that ensure that the procedure removes the
relationship between sampling depth and uncorrected variance and covariances
(Supplementary Figure \ref{suppfig:bergland-correction}).  \vb{Through our
  simulations we find that our estimates can differ based on how fixations and
  losses are handled in long time-series (Supplementary Material Section
  \ref{supp:fixation}) but none of our findings in the main text are altered by
  this decision (Supplementary Material Figures \ref{suppfig:supp-fig-1-nofix}
  and \ref{suppfig:supp-fig-3-nofix}).}

@@

\subsection{Approximating the Reduction in Diversity from $G(t)$}

To help help reconcile our measure of linked selection, $G(t)$, with familiar
reductions  \gc{in neutral diversity, as parameterized by} $N_e$, we develop some rough
approximations here. First, we introduce some convenient notation. Let $V_{T} =
\var(p_t - p_0)$ be the total observed variation in allele frequency, $C_{LS} =
\sum_{i=0}^{t-1} \sum_{j \ne i}^{t-1} \cov(\Delta p_i, \Delta p_j)$ be the
contribution of all pairwise temporal covariances (observable), $V_D$ be the
unobservable drift-only variance in allele frequency, and $V_{LS} =
\sum_{i=0}^{t-1} \var(\Delta_{LS} p_i$ which is the (unobservable) effect
linked selection has on the variances in allele frequency change.

\begin{align}
  V_{T} = V_{LS} + C_{LS} + V_D
\end{align}

Our measure $G(t)$ is then,

\begin{align}
  G(t) = \frac{C_{LS}}{V_T}
\end{align}

meaning we can express the fraction of total variance due to drift alone as

\begin{align}
  \frac{V_D}{V_T} &= 1 - G(t) - \frac{V_{LS}}{V_T} \\
                  &= 1 - G(t) - \varepsilon \\
                  & \le 1 - G(t).
\end{align}

where $\varepsilon \ge 0$ since these variances must be positive.

\gc{In a neutral} Wright--Fisher population,

\begin{align}
  V_T = \var(p_t - p_0) = p_0(1 - p_0) \left[ 1 - (1 - \nicefrac{1}{2N_e})^t \right]
\end{align}

For small $t$, a common temporal estimator for the variance effective
population size $N_e$ is,

\begin{align}
  G(t) \lesssim 1 - \frac{N_e}{N_{e,D}} 
\end{align}
which shows that our measure $G(t)$ is a lower bound, over much shorter
timescales, to the familiar measure $1 -
\nicefrac{\bar{\pi}}{\pi_{\gc{0}}}$.

\gc{\cite{Elyashiv2016-vt} estimated that linked selection in \emph{Drosophila
  melanogaster} had resulted in a $1 -
\nicefrac{\bar{\pi}}{\pi_{\gc{0}}} = 77\%$ reduction in the level of
neutral diversity. Thus our estimate of $G(t) \approx 20\%$ in \emph{Drosophila
  simulans} is smaller
than that seen over long timespans in a closely related
species. }

\subsection{Forward-in-time Simulations}
\label{supp:forward}
\vb{
To explore how aspects of genetic architecture, model of selection, and
experimental design impact temporal covariance, the $G(t)$ trajectories, and
convergence correlations, we ran extensive forward-in-time using SLiM
\parencite{Haller2019-vu}; here we discuss the Gaussian Stabilizing Selection
simulations in Figured \ref{fig:figure-4}, but Supplementary Materials Section
\ref{supp:forward} describes these simulation routines in much more detail as
well as additional simulations we have conducted.

We simulated directional selection on a trait by first burning in a population
of $N = 1000$ diploids under GSS for $10N$ generations with the stabilizing
selection trait variance $V_s = 1$ and an optima set at zero. We targeted a
polygenic architecture by setting the trait mutation rate to $10^{-8}$ per
basepair, per generation, in addition to having a separate neutral mutation of
$10^{-8}$ which created neutral mutations which we used to calculate the
temporal covariances. Our simulated region was 50 megabases in length (about
one quarter of a \emph{Drosophila} chromosome), and trait alleles were randomly
selected to have a $\pm 0.01$ effect size. By tracking the trait mean through
the burnin, we found it converged as expected. After a $10N$ burnin, the
population was split into two different replicate population sizes, to capture
the effect of bottlenecks in selection experiments (these population sizes
were, 50, 500, and 1000 diploids; the later representing no bottleneck). Each
population then underwent an optima shift of either 0.1, 0.5, or 1 on
generation five, with the first four generations serving as a control. These
optima shifts were either in the same direction (converging), different
directions (diverging), or only one optima shifted (as a control). By tracking
the trait mean, we saw that it converged as expected during burnin, and had the
appropriate response to selection (Supplementary Material Figure
\ref{suppfig:gss-zbar}). Using the neutral population frequency data from these
simulations, we calculated the temporal covariances, $G(t)$ trajectories, and
convergence correlations.
}

@@

>>>>

 Rapid phenotypic adaptation is often observed in natural populations and
 selection experiments.  However, detecting the genome-wide impact of this
 selection is difficult, since adaptation often proceeds from standing
 variation and selection on highly polygenic traits, both of which may leave
 faint genomic signals indistinguishable from a noisy background of genetic
 drift.  One promising signal comes from the genome-wide covariance between
 allele frequency changes observable from temporal genomic data, e.g.
 evolve-and-resequence studies. These temporal covariances reflect how
   heritable fitness variation in the population leads the changes in allele
   frequencies at one timepoint to be predictive of the changes at later
 timepoints, as alleles are indirectly selected \gc{due} to remaining
 associations with selected \gc{alleles}. Since genetic drift does not lead to
 temporal covariance, we can use these covariances to estimate what fraction of
 the variation in allele frequency change through time is driven by linked
 selection. Here, we reanalyze \vb{three selection experiments} to quantify the
 effects of linked selection over short timescales using covariance among
 time-points and across replicates. We estimate that at least 17\% to 37\% of
 allele frequency change is driven by selection in these experiments.  Against
 this background of positive genome-wide temporal covariances we also identify
 signals of negative temporal covariance corresponding to reversals in the
 direction of selection for a reasonable proportion of loci over the time
 course of a selection experiment.  Overall, we find that in the three studies
 we analyzed, linked selection has a large impact on short-term allele
 frequency dynamics that is readily distinguishable from genetic drift.

@@

A long-standing problem in evolutionary genetics is quantifying the roles of
genetic drift and selection in shaping genome-wide allele frequency changes.
Selection can affect allele frequencies, \vb{both directly and indirectly,}
with the indirect effect coming from the action of selection on \gc{correlated} %\vb{linked}
loci elsewhere in genome e.g. linked selection (\cite{Maynard_Smith1974-lc},
\cite{Charlesworth1993-gb,Nordborg1996-nq}; see \cite{Barton2000-zg} for a
review). Previous work \vbout{on this question} has mostly focused on teasing
apart the impacts of drift and selection on genome-wide diversity using
population samples from a single contemporary timepoint, often by modeling the
correlation between regional recombination rate, gene density, and diversity
created in the presence of linked selection
\parencite{Cutter2013-ba,Sella2009-nx}. This approach has shown linked
selection has a major role in shaping patterns of genome-wide diversity across
the genomes of a range of sexual species
\parencite{Macpherson2007-qt,Andolfatto2007-uy,Begun2007-bg,Beissinger2016-cm,Sattath2011-dr,Williamson2014-oy,Andersen2012-bj,Cutter2010-gi,Elyashiv2016-vt},
and has allowed us to quantify the relative influence of positive selection
(hitchhiking) and negative selection (background selection;
\cite{Nordborg2005-dc,McVicker2009-ax,Andolfatto2007-uy,Macpherson2007-qt,Hernandez2011-gs,Elyashiv2016-vt}).
However, we lack an understanding of both how linked selection acts over
\vb{short time intervals and its full impact on genome-wide allele frequency
changes.}

There are numerous examples of rapid phenotypic adaptation
\parencite{Grant2011-wk,Grant2006-hj,Reznick1997-mh,Franks2007-dr} and rapid,
selection-driven genomic evolution in asexual populations
\parencite{Good2017-om,Bennett1990-bc,Baym2016-kh}.  Yet the polygenic nature
of fitness makes detecting the impact of selection on genome-wide variation
over short timescales in sexual populations remarkably difficult
\vb{\parencite{Latta1998-me, Pritchard2010-tk,Kemper2014-bx}}. This is because
the effect of selection on a polygenic trait (such as fitness) is distributed
across \vb{numerous} loci. This can lead to subtle allele frequency shifts on
standing variation that are difficult to distinguish from background levels of
genetic drift and sampling variance. \vb{Increasingly}, genomic experimental
evolution studies with multiple timepoints, and in some cases multiple
replicate populations, are being used to detect large effect selected loci
\parencite{Turner2011-sx,Turner2012-bm} and differentiate modes of selection
\parencite{Burke2010-tz,Barghi2019-qy,Therkildsen2019-zy}.  In addition these
temporal-genomic studies have begun in wild populations, some with the goal of
finding variants that exhibit frequency changes consistent with fluctuating
selection \parencite{Bergland2014-ij,Machado2018-cs}. In a previous paper, we
proposed that one useful signal for understanding the genome-wide impact of
polygenic linked selection detectable from genomic studies with multiple
timepoints is the temporal autocovariance (\vb{i.e. covariance between two
timepoints}) in allele frequency changes \parencite{Buffalo2019-io}.  These
covariances \vb{can be estimated} from temporal genomic data and are created
when the loci that underly heritable fitness variation perturb the frequencies
of linked alleles; in contrast, when genetic drift acts alone in a closed
population, these covariances are zero in expectation for neutral alleles.
Mathematically, temporal covariances are useful because it is natural to
decompose the total variance in allele frequency change across a set of time
intervals into the variances and covariances in allele frequency change among
time intervals.  Furthermore, biologically, these covariances reflect the
extent to which allele frequency changes in one generation predict changes in
another due to a shared selection pressures and associations to selected loci.

@@

We first analyzed \textcite{Barghi2019-qy}, an evolve-and-resequence study with
ten replicate populations exposed to a high temperature lab environment and
evolved for 60 generations, and sequenced every ten generations. Using the
seven timepoints and ten replicate populations, we estimated the genome-wide $6
\times 6$ temporal covariance matrix $\mathbf{Q}$ for each of the ten
replicates. Each row of these matrices represent the temporal covariance
$\cov(\Delta_{_{10}} p_s, \Delta_{_{10}} p_t)$, between the allele frequency
change (in ten-generation intervals, denoted $\Delta_{_{10}} p_t$) in some
initial reference generation $s$ (the row of the matrix), and some later
timepoint $t$ (the column of the matrix). We corrected these matrices for
biases created due to sampling noise, and normalize the entries for
heterozygosity (see Supplementary Materials Sections
\ref{supp:ind-depth-var-corr} and \ref{supp:matrix-correction}). These
covariances are expected to be zero when only drift is acting, as only
heritable variation for fitness can create covariance between allele frequency
changes in a closed population \parencite{Buffalo2019-io}.  Averaging across
the ten replicate temporal covariances matrices, we find temporal covariances
that are statistically significant (95\% block bootstraps CIs do not contain
zero), consistent with linked selection perturbing genome-wide allele frequency
changes over very short time periods. The covariances between all adjacent time
intervals are positive and then decay towards zero as we look at more distant
time intervals (Figure \ref{fig:figure-1} A), as expected when directional
selection affects linked variants' frequency trajectories until ultimately
linkage disequilibrium and the associated additive genetic variance for fitness
decays (\vb{which could occur as a population reaches a new optimum})
\parencite{Buffalo2019-io}. The temporal covariances per replicate are noisier
but this general pattern holds; see Supplementary Figure
\ref{suppfig:barghi-cov-panels}.  \textcite{Barghi2019-qy}'s design means that
the covariances we see in adjacent time intervals are on average ten
generations apart, and given the temporal decay in covariance we see, the
covariances on shorter time-scales (e.g. if adjacent generations had been
sequenced) may well be higher yet (see Supplementary Material Section
\ref{supp:barghi-covs} for more details).

@@

which estimates the \vb{impact of selection} on allele frequency change between
the initial generation $0$ and some later generation $t$, which can be varied
to see how this quantity grows through time. \vb{This measure can intuitively
  be understood as the relative fraction of allele frequency change normally
  thought of as ``drift" that is actually due to selection; the denominator of
  $G(t)$ is $\var(p_t - p_0)$, which under drift alone is proportional to
  $\nicefrac{1}{2 N_e}$. By measuring the contribution of temporal covariances
caused by selection, $G(t)$ looks at the fraction selection contributes to what
looks like a faster rate of drift, or lower $N_e$. Additionally, $G(t)$ can be
understood as a short-timescale estimate of the reduction in neutral diversity
due to linked selection (see Supplementary Materials Section
\ref{supp:reduction-approx}).} Since \textcite{Barghi2019-qy} experiment is
sequenced every ten generations, in the numerator for the covariance we use the
allele frequency changes between adjacent timepoints, which are ten generations
apart; \vb{thus the strong covariances between adjacent generations are not
  observable and do not contribute to the numerator of $G(t)$}.
  \vb{Additionally, selection also inflates the variance in allele frequency
    change per generation; however, this effect cannot be easily distinguished
    from drift.} \vb{For both these reasons, our measure $G(t)$ is strongly
  conservative} (\vb{we demonstrate this through simulations in Supplementary
Material Section \ref{supp:tempblocks}}).  Still, we find a remarkably strong
signal. Greater than $20\%$ of total, genome-wide allele frequency change over
60 generations is the result of selection (Figure \ref{fig:figure-1} B).

@@

The replicate design of \textcite{Barghi2019-qy} allows us to quantify another
covariance: the covariance in allele frequency change between replicate
populations experiencing convergent selection pressures. These
between-replicate covariances are created in the same way as temporal
covariances: alleles linked to a particular fitness background are
expected to have allele frequency changes in the same direction if the
selection pressures are similar. Intuitively, where temporal covariances
reflect that alleles associated with heritable fitness backgrounds are
predictive of frequency changes between generations, replicate covariances
reflect that heritable fitness backgrounds common to each replicate predict
(under the same selection pressures) frequency changes between replicates. We
measure this through a statistic similar to a correlation, which we call the
convergent correlation: the ratio of average between-replicate covariance
across all pairs to the average standard deviation across all pairs of
replicates,

@@

Our primary focus here has been on evolution in laboratory populations. It is
unclear whether we should expect a similar impact of selection in natural
populations. In some of these experiments, selection pressures may have been
stronger or more sustained \vb{than} in natural populations
\parencite{Hendry1999-zu,Hairston2005-ga}. Conversely, these lab populations
were maintained at relatively small census sizes (Table
\ref{table:study_summary}), which will amplify the role of genetic drift,
\vb{and increase the frequency of rare deleterious alleles in selection lines
due to founder effects}. The advantage of lab experiments is that they are
closed populations\vb{;} in natural populations temporal covariance could also
arise from the systematic migration of alleles from differentiated populations.
Adapting these methods to natural populations will require either populations
that are reasonably closed to migration, or for the effect of migration to be
accounted for possibly either by knowledge of allele frequencies in source
populations or the identification of migrant individuals.

@@

To remove systematic covariances in allele frequency change caused by tracking
the reference or minor allele, we randomly choose which allele's frequency to
track for each locus. Then, we calculate the variance-covariance matrix of
allele frequency changes using a Python software package we have written,
available at \url{http://github.com/vsbuffalo/cvtk}. This simultaneously
calculates temporal variances and covariances, and replicate covariances and
uses the sampling depth and number of diploid individuals to correct for bias
in the variance estimates and a bias that occurs in covariance estimates
between adjacent timepoints due to shared sampling noise (see Supplementary
Material Sections \ref{supp:ind-depth-var-corr}, \ref{supp:cov-corr}, and
\ref{supp:matrix-correction} for mathematical details of these estimators). We
assess that our bias correction procedure is working adequately through a
series of diagnostic plots that ensure that the procedure removes the
relationship between sampling depth and uncorrected variance and covariances
(Supplementary Figure \ref{suppfig:bergland-correction}).  Through our
  simulations we find that our estimates can differ based on how fixations and
  losses are handled in long time-series (Supplementary Material Section
  \ref{supp:fixation}) but none of our findings in the main text are
  \gc{qualitatively} altered by
  this decision (Supplementary Material Figures \ref{suppfig:supp-fig-1-nofix}
  and \ref{suppfig:supp-fig-3-nofix}).

@@

\subsection{Forward-in-time Simulations}
\label{supp:forward}

To explore how aspects of genetic architecture, model of selection, and
experimental design impact temporal covariance, the $G(t)$ trajectories, and
convergence correlations, we ran extensive forward-in-time using SLiM
\parencite{Haller2019-vu}; here we discuss the Gaussian Stabilizing Selection
simulations in Figured \ref{fig:figure-4}, but Supplementary Materials Section
\ref{supp:forward} describes these simulation routines in \gc{more} detail as
well as additional simulations we have conducted.

We simulated directional selection on a trait by first burning in a population
of $N = 1000$ diploids under GSS for $10N$ generations with the stabilizing
selection trait variance $V_s = 1$ and an optima set at zero.
\gc{We note that the small burnin population size means that these
  simulations should not be taken as reflecting any specific natural
  population and they are for illustrative purposes only.}
We targeted a
polygenic architecture by setting the trait mutation rate to $10^{-8}$ per
basepair, per generation, in addition to having a separate neutral mutation of
$10^{-8}$ which created neutral mutations which we used to calculate the
temporal covariances. Our simulated region was 50 megabases in length (about
one quarter of a \emph{Drosophila} chromosome), and trait alleles were randomly
selected to have a $\pm 0.01$ effect size. By tracking the trait mean through
the burnin, we found it converged as expected. After a $10N$ burnin, the
population was split into two different replicate population sizes, to capture
the effect of bottlenecks in selection experiments (these population sizes
were, 50, 500, and 1000 diploids; the later representing no bottleneck). Each
population then underwent an optima shift of either 0.1, 0.5, or 1 on
generation five, with the first four generations serving as a control. These
optima shifts were either in the same direction (converging), different
directions (diverging), or only one optima shifted (as a control). By tracking
the trait mean, we saw that it converged as expected during burnin, and had the
appropriate response to selection (Supplementary Material Figure
\ref{suppfig:gss-zbar}). Using the neutral population frequency data from these
simulations, we calculated the temporal covariances, $G(t)$ trajectories, and
convergence correlations.

@@

\subsection{Approximating the Reduction in Diversity from $G(t)$}
\label{supp:reduction-approx}

To help reconcile our measure of linked selection, $G(t)$, with familiar
\vb{expressions as a reduction} in neutral diversity, as parameterized by
$N_e$, we develop some rough approximations here. First, we introduce some
convenient notation. Let $V_{T} = \var(p_t - p_0)$ be the total observed
variation in allele frequency, $C_{LS} = \sum_{i=0}^{t-1} \sum_{j \ne i}^{t-1}
\cov(\Delta p_i, \Delta p_j)$ be the contribution of all pairwise temporal
covariances (observable), $V_D$ be the unobservable drift-only variance in
allele frequency, and $V_{LS} = \sum_{i=0}^{t-1} \var(\Delta_{LS} p_i)$ which is
the (unobservable) effect linked selection has on the variances in allele
frequency change. Then,

\begin{align}
  V_{T} = V_{LS} + C_{LS} + V_D
\end{align}
%
Our measure $G(t)$ is then,

\begin{align}
  G(t) = \frac{C_{LS}}{V_T}
\end{align}
%
meaning we can express the fraction of total variance due to drift alone as

\begin{align}
  \frac{V_D}{V_T} &= 1 - G(t) - \frac{V_{LS}}{V_T} \\
                  &= 1 - G(t) - \varepsilon \\
                  & \le 1 - G(t),
\end{align}
%
where $\varepsilon \ge 0$ since these variances are positive. \vb{Throughout
this section, for convenience, we assume that the covariances contributing to
$C_{LS}$, and thus $G(t)$, are all positive.}

In a neutral Wright--Fisher population, the total variation in allele
frequency change over $t$ generations is

\begin{align}
  V_T = \var(p_t - p_0) = p_0(1 - p_0) \left[ 1 - (1 - \nicefrac{1}{2N_e})^t \right].
\end{align}
%
For small $t$, a common temporal estimator for the variance effective
population size $N_e$ is,

\begin{align}
  G(t) \lesssim 1 - \frac{N_e}{N_{e,D}} 
\end{align}
which shows that our measure $G(t)$ is a lower bound, over much shorter
timescales, to the familiar measure $1 -
\nicefrac{\bar{\pi}}{\pi_{0}}$.

\textcite{Elyashiv2016-vt} estimated that linked selection in \emph{Drosophila
  melanogaster} had resulted in a $1 - \nicefrac{\bar{\pi}}{\pi_{0}} = 77\%$
  reduction in the level of neutral diversity. Thus our estimate of $G(t)
  \approx 20\%$ in \emph{Drosophila simulans} is smaller than that seen over
long timespans in a closely related species. }
